---
title: "Intro to readxl"
author: "Sean McKnight"
date: "December 7, 2019"
output: html_document
---

# Learning `readxl`
### Hello!

We are working with data in readxl and it's going to be amazing

We're going to use packages: 

- tidyverse
- readxl
- writexl
- here

```{r packages}
library(tidyverse)
library(readxl)
library(writexl)
library(here)

# this will attach the package i.e., turn the lights on
# first need to do install.packages("package_name")
```

### Read in fish counts

```{r}
# read in fish_counts_curated.csv
fish_counts <-read_csv(here("R-workshop-data","fish_counts_curated.csv"))
```

### Read in data from a single Excel worksheet

Now, to read in an excel spreadsheet, we'll use the function 'read_excel()'.

```{r}
substrate_cover <- read_excel(here("R-workshop-data","substrate_cover_curated.xlsx"),
                              skip = 4,
                              na = "-9999")
```


### Use read_excel() to read in a range of values

```{r}
carp_cover_2000 <- read_excel(here("R-workshop-data",
                                   "substrate_cover_curated.xlsx"), range = "D12:I15",
                              col_names = c("site",
                                            "transesct",
                                            "quad",
                                            "side",
                                            "type",
                                            "coverage"))
```


### How can we read in data from an Excel workbook with MULTIPLE Worksheets?


```{r}
kelp_golb <- read_excel(here("R-workshop-data",
                             "kelp_counts_curated.xlsx"),
                        sheet = "golb")
```

### Read in AND combine data from multiple Excel worksheets simultaneously using purr::map_df()

```{r}
# create a path to get the data
# kelp_path <- here("R-workshop-data","kelp_counts_Curated.xlsx")
# 
# kelp_all_sites <- kelp_path %>%
#   excel_sheets() %>% 
#   set_names() %>% 
#   map_df (read_excel, kelp_path)
                  

```


### Write a file back to an excel spreadsheet

We're going to write our "Kelp_golb" object to a CSV file.

```{r}
write_csv(kelp_golb, "kelp_goleta_july.csv")
#alternatively, you can use the here() option 
#write_csv(kelp_golb, here())

write_xlsx(kelp_golb,"kelp_goleta_excel.xlsx")
```

test text text test
