---
title: "Filter/joins (like VLOOKUP in Excel)"
author: "Sean McKnight"
date: "12/8/2019"
output: html_document
---

## Attach Packages:

tidyverse, readxl, here

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Read in 3 files: 

- invert_counts_curated.xlsx
- fish_counts_curated.csv
- kelp_counts_curated.xlsx

```{r}
invert_counts <- read_xlsx(here("R-workshop-data/invert_counts_curated.xlsx" ))
fish_counts <- read_csv(here("R-workshop-data/fish_counts_curated.csv" ))
kelp_counts_abur <- read_xlsx(here("R-workshop-data/kelp_counts_curated.xlsx" ))
```


Notes: select vs filter:
- select() function: keep, exclude or rearrange columns
- filter() function: keeps or excludes ROWS (observations) basedon whether or no they match variable conditions

## Filter to match rows based on character string

Only keep observations from fish_counts where the common_name variable matches "garibaldi"


```{r}
fish_garibaldi <- fish_counts %>%
  filter(common_name == "garibaldi") #in R, you double = when searching for something

# use janitor::clean_names() to automatically update all of your column names to lowercase_snake_case
```


##Filter to keep/exclude observations based on numeric conditions

create a subset where total_counts is >=50

```{r}
fish_over50 <- fish_counts %>%
  filter(total_count >= 50)
```


Create a subset that only contains observations for fish at Mohawk Reef ("mohk"):
```{r}
fish_mohk <- fish_counts %>%
  filter(site == "mohk")
```


## Filter to keep rows that match THIS OR THAT

Example: We want tot keep any observations for garibaldi, blacksmith, or black surfperch

Use the %in% operator to look for any matches within a vector c() is a vector

```{r}
# this is verymuch an OR statement
fish_3sp <- fish_counts %>% 
  filter(common_name %in% c("garibaldi","blacksmith","black surfperch"))

# Writing OR statements explicity:
# Use the vertical line | to indicate "or"
fish_3sp_v2 <- fish_counts %>%
  filter(common_name == "garibaldi" | common_name == "blacksmith" | common_name == "black surfperch")
```


## Filter to return rows that match conditions for MULTIPLE variables (AND statements)

I want to create a subset called urchin_abur_mohk that only kees observations where: 

- site is "abur" or "mohk" AND
- species is "purple urchin"

Some ways to do this: 

- separate "AND" conditions within a signle filter line by a comma
- separate "AND" conditions within a single filter line by an ampersand (&)
- putting each filter condition within a separate filter line that are just piped together in sequence

```{r}
urchin_abur_mohk <- invert_counts %>%
  filter(site %in% c("abur","mohk"), common_name == "purple urchin")

urchin_abur_mohk2 <- invert_counts %>%
  filter(site %in% c("abur","mohk") & common_name == "purple urchin")

```


