---
title: "intro-to-ggplot2"
author: "Sean McKnight"
date: "12/8/2019"
output: html_document
---

##Attach packages
```{r}
library(tidyverse)
library(readxl)
library(here)
```

##Read in the data
```{r}
#use here() to indicate subfolder
ca_np <- read_csv(here("R-workshop-data","ca_np.csv"))

ci_np <- read_xlsx(here("R-workshop-data","ci_np.xlsx"))
```


## Make a graph w/ ggplot2

Need three things:

1. Tell R wer're using `ggplot`
2. Tell R what data we're using, and what variables we want to plot
3. What type of graph we're creating (geom)

```{r, eval = FALSE}
#eval = FALSE means the code won't run
#these are the basics of having a minimal viable graph

ggplot(data = THIS, aes(x = THIS, y = THIS)) + 
  geom_THIS()
```

Now let's create a graph of CI visitation:
```{r}
#aes = aesthetic
#for ggplot, use the vertical structure with plus signs, the plus sign adds indents
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

```

## try a scater plot

```{r}
#copy from above and change geom
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```

## Create a column graph with this data

```{r}
#col = column / bar chart
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()
```

## Create an area graph with this data
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```


## Intro to customozing  `ggplot` graphs
- color = (update point or line color)
- fill = (update the fill characteristics of anelement)
- linetype =
- pch = (value)
- size = 
- alpha = 

```{r}
#change elements of the graph within geom
#only change aes are only the variables. 
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(color = "coral4",
            linetype = "dashed")
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point (color = "blue",
            pch = 17,
            size = 4,
            alpha = 0.5)
```

I have some awesome text here, and below is my amazing final graph:
```{r, echo = FALSE}
#alpha = transparancy, values between 0 and 1. Closer to 0 = more transparent, closer to 1 = less transparent
# echo = FALSE is hiding the code in the markdown
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point (color = "blue",
            pch = 18,
            size = 4,
            alpha = 0.5)
```

