---
title: "intro-to-ggplot2"
author: "Sean McKnight"
date: "12/8/2019"
output: html_document
---

##Attach packages
```{r}
library(tidyverse)
library(readxl)
library(here)
```

##Read in the data
```{r}
#use here() to indicate subfolder
ca_np <- read_csv(here("R-workshop-data","ca_np.csv"))

ci_np <- read_xlsx(here("R-workshop-data","ci_np.xlsx"))
```


## Make a graph w/ ggplot2

Need three things:

1. Tell R wer're using `ggplot`
2. Tell R what data we're using, and what variables we want to plot
3. What type of graph we're creating (geom)

```{r, eval = FALSE}
#eval = FALSE means the code won't run
#these are the basics of having a minimal viable graph

ggplot(data = THIS, aes(x = THIS, y = THIS)) + 
  geom_THIS()
```

Now let's create a graph of CI visitation:
```{r}
#aes = aesthetic
#for ggplot, use the vertical structure with plus signs, the plus sign adds indents
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

```

## try a scater plot

```{r}
#copy from above and change geom
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```

## Create a column graph with this data

```{r}
#col = column / bar chart
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()
```

## Create an area graph with this data
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```


## Intro to customozing  `ggplot` graphs
- color = (update point or line color)
- fill = (update the fill characteristics of anelement)
- linetype =
- pch = (value)
- size = 
- alpha = 

```{r}
#change elements of the graph within geom
#only change aes are only the variables. 
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(color = "coral4",
            linetype = "dashed")
```

```{r}
#pch = symbol type
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point (color = "blue",
            pch = 17,
            size = 4,
            alpha = 0.5)
```

I have some awesome text here, and below is my amazing final graph:
```{r, echo = FALSE}
#alpha = transparancy, values between 0 and 1. Closer to 0 = more transparent, closer to 1 = less transparent
# echo = FALSE is hiding the code in the markdown
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point (color = "blue",
            pch = 18,
            size = 4,
            alpha = 0.5)
```


## Customize elements based on variable values or outcomes

```{r}
# create bubble plot (look cool, but are often just confusing)
# if your aes is based on a variable, you have to embed that in the geom 
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point(aes(size = visitors,
                 color = visitors),
             pch = 17,
             alpha = 0.8) 
```

For this activity, create a column plot of CINP visitaiton over time, where the fill color changes based on the number of visitors, and the outline color (color = ) is "magenta"

```{r}
# the fill color is based on a variable, so it is inside aes, the outline color is outside the aes b/c it's not based on a variable.
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col(aes(fill = visitors),
           color = "magenta")

```


```{r}
# The theme_THIS is how you adjust the background, you can create your own specific theme like theme_sean and apply that any time you make a graph
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(size = 3,
            color = "red") +
  theme_minimal()
```


```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point(size = 1,
            color = "navy") +
  theme_minimal() +
  labs (x = "My x-axis title",
        y = "Annual visotors",
        title = "Channel Islands NP Visitation",
        subtitle = "(1963 - 2016")
```

## Combine compatible geoms

Example: Add a point plot on top of a line plot: 

```{r}
# whatever comes first will be farthest back
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(color = "purple") +
  geom_point(color = "magenta",
             aes(size = year)) +
  theme_minimal()
```


Create a combined column graph + line graph of visitation. 

```{r}
# 
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()+
  geom_line(color = "navy",
            size = 2,
            alpha = 0.8) +
  theme_minimal()
```


##Create a line graph with multiple series

```{r}
#color = park_name is another variable, and R knows that there are variables.
ggplot(data = ca_np, aes(x = year, y = visitors, color = park_name)) +
  geom_line()
```

## Faceting ggplot graphs

- each series (group) gets its own plot panel

```{r}
# the facet wrap
# show 
ggplot(data = ca_np, aes(x = year, y = visitors, color = park_name)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ park_name)+
  theme_minimal()

```

## Use ggsave to export a plot

```{r}
#this will save the most recent graph, you cold save call a specific graph
ggsave("np_graph.png", dpi = 180, width = 8, height = 7)
```

