---
title: "Pivot tables"
author: "Sean McKnight"
date: "12/8/2019"
output: html_document
---

# Restart R (session > restart)
# Attach libraries
```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr) #install.packages("skimr")

```


##Read in our data
our data is about lobsters
```{r}

lobsters <- read_xlsx(here("R-workshop-data/lobsters2.xlsx" ))
```

Explore our data
```{r}
#head() function will show the first 6 lines of data
head(lobsters)

#summary() to give a summary of each column, look for NA's and know what kind of beast you are dealing with.
summary(lobsters)

#skimr::skim() gives additional summary stats and is pipalbe (%>%)
skim(lobsters)

# ggplot of lobsters
ggplot(data = lobsters, aes(x = year)) +
  geom_bar() +
  facet_wrap(~site)
```


## Our Task: 

How has the average size of lobsters has changed over time across sites?

Pivot tables in R are group_by %>% sumarize (command shift M for piping shortcut)
```{r}
lobsters %>% #in your head wwhen you see %>%  think "and then" helps you unnest something, and pass the data to the next argument. Think of it as story telling. Take the lobsters data and then group it by year, and then summarize count by year
  group_by(year) %>% 
  summarize(count_by_year = n())

```


What if we didnt' group_by first?
```{r}
lobsters %>%
  summarize(count_by_something = n())

## assign our summary to a new variable
lobsters_sum <- lobsters %>%
  summarize(count_by_something = n())
  
```

## group_by multiple variables

and summarize by multiple variables

We used mean() to calculate the mean size of lobsters by earch site and year, and used na.rm = TRUE to calculate mean despite having missing values in our raw data
```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))

siteyear_summary
```

# Activity

1. Calculate the median size_mm (?median)
2. Create and ggsave() a plot

```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))

siteyear_summary

##ggplot: geom_line
ggplot(data = siteyear_summary, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()

# I created a figurs folder form the RStudio Files pane in order to be organized and for this code to run
ggsave(here("figures/lobsters-line.png"))

```


# dplyr::coutn() for county by group

```{r}

lobsters %>% 
  count(site, year)

## this is the same as

lobsters %>%
  group_by(site, year) %>% 
  summarize(n = n()) %>% 
  ungroup()
```


# mutate() add a new column
it will make a new variable

- Add a new column that is the size in meters
- Add a new column that is a repeated value

```{r}
head(lobsters)

lobsters %>% 
  mutate(size_m = size_mm / 1000,
         millenia = 2000,
         observer = "Allison Horst")

```


## Select() function to subset variables by columns

let's select 
- date
- site
- size_m

```{r}

lobsters_selected <- lobsters %>%
  mutate(size_m = size_mm / 1000) %>% 
  select(date, site, size_m)

```


